Optimizing the future of biodiversity sampling by citizen scientists

### 2019-04-10 paper/main.tex
added a 'methods' para before results

Here, we investigate these questions with a specific objective: investigate how spatial and temporal sampling influences trend detection of biodiversity. Our approach is site-specific and dynamic: we are interested in the parameters that influence the marginal value of a given citizen science observation for a particular day. We test our approach using a popular citizen science project --- eBird \cite{sullivan2009ebird} --- and > 5 million bird biodiversity observations from the Greater Sydney Region, NSW, Australia. First, we summarize the parameters in space and time which we hyptohesized would influence the marginal value of a citizen science observation.

@@

>>>>

Here, we investigate these questions with a specific objective: investigate how spatial and temporal sampling influences trend detection of biodiversity. Our approach is site-specific and dynamic: we are interested in the parameters that influence the marginal value of a given citizen science observation for a particular day. First, we summarize the parameters in space and time which we hyptohesized would influence the marginal value of a citizen science observation.

\subsection*{predictions}

@@

We test our approach using a popular citizen science project --- eBird \cite{sullivan2009ebird} --- and > 5 million bird biodiversity observations from the Greater Sydney Region, NSW, Australia. We fit presence/absence population trend models to 235 species throughout this region, based on data from 2010 to 2018 (Fig. 1), and calculate statistical leverage for every citizen science sampling event, for each species. The leverage is a measure of how 'valuable' a given observation is to the population trend model for each species. Each sampling event is then assigned a cumulative measure of leverage --- i.e., marginal value. We then test our predictions above using linear models, assessing the influence each of the parameters has on the trend detection of biodiversity. Finally, we use the parameterized model results to predict the expected leverage (i.e., marginal value) for a given grid (i.e., site) in space and time, deriving a dynamic map of expected marginal value for every day in 2018.
